---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
First, I will unzip the files using the unzip function. Then, I will read in the resulting csv into the dataframe titled activity, using the read.csv function.
I also ran a str function to see, exactly, which variables are present within the dataset.
```{r unzip, echo=TRUE}
unzip("activity.zip")
activity<-read.csv("activity.csv")
str(activity)
```


## What is mean total number of steps taken per day?
```{r mean, echo=TRUE}
library("dplyr")
#calculate total steps per day 
stepsbyday<-activity %>% select(steps, date) %>% group_by(date) %>% summarize(stepsbyday=sum(steps, na.rm=TRUE))
#create histogram of total steps per day 
hist(stepsbyday$stepsbyday)
#calculate mean and median steps per day 
mean(stepsbyday$stepsbyday, na.rm=TRUE)
median(stepsbyday$stepsbyday, na.rm=TRUE)
```


## What is the average daily activity pattern?
```{r average, echo=TRUE}
library("dplyr")
library("ggplot2")
#calculate average number of steps taken over the 5-minute intervals 
stepsbyday<-activity %>% select (steps, date, interval) %>% group_by(interval) %>% summarize(stepsbyday=mean(steps, na.rm=TRUE))
#graph the plot
ggplot(stepsbyday, aes(x=interval, y=stepsbyday))+ geom_line()+labs(title="avg.steps daily in intervals")
#calculate interval that has max number of steps 
stepsbyday[stepsbyday$stepsbyday==max(stepsbyday$stepsbyday),]
```


## Imputing missing values
According to the xray package,there is approximately 13.11% missing. I subsetted the missing steps, placed them into a dataframe named ismissing, and then used dplyr to spell out the specific dates with missing steps. 
```{r whichismissing, echo=TRUE}
library("xray")
anomalies(activity)
ismissing<-(activity[is.na(activity$steps),])
ismissing %>% distinct(date)
```
To impute missing values, I am going to use the "roughfix" portion of the randomForest package. Since steps is numeric, it imputes the median for the variable. As we can see, the median of both of the dataframes are equal. While the means have a 5 step gap, the variance are quite close. 

Compared to the estimate, the mean is approximately 2-3K steps higher. Evidently, imputing the missing value has an impact. 

```{r imputetime, echo=TRUE}
library("randomForest")
set.seed(20)
ismissing.roughfix<-na.roughfix(activity)
median(activity$steps, na.rm=TRUE)
median(ismissing.roughfix$steps)

mean(activity$steps,na.rm=TRUE)
mean(ismissing.roughfix$steps)

var(activity$steps, na.rm=TRUE)
var(ismissing.roughfix$steps)
```

```{r histogram, echo=TRUE}
#recalculate total steps per day 
stepsbyday<-ismissing.roughfix %>% select(steps, date) %>% group_by(date) %>% summarize(stepsbyday=sum(steps, na.rm=TRUE))
ggplot(stepsbyday, aes(x=stepsbyday))+geom_histogram()+labs(title="total steps")
```
## Are there differences in activity patterns between weekdays and weekends?

```{r creatingweekdaysandweekends, echo=TRUE}
#ismissing.roughfix$date<-as.Date(ismissing.roughfix$date)
ismissing.roughfix$date<-weekdays(ismissing.roughfix$date)
```
