---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
First, I will unzip the files using the unzip function. Then, I will read in the resulting csv into the dataframe titled activity, using the read.csv function.
I also ran a str function to see, exactly, which variables are present within the dataset.
```{r unzip, echo=TRUE}
unzip("activity.zip")
activity<-read.csv("activity.csv")
str(activity)
```


## What is mean total number of steps taken per day?
```{r mean, echo=TRUE}
library("dplyr")
#calculate total steps per day 
stepsbyday<-activity %>% select(steps, date) %>% group_by(date) %>% summarize(stepsbyday=sum(steps, na.rm=TRUE))
#create histogram of total steps per day 
hist(stepsbyday$stepsbyday)
#calculate mean and median steps per day 
mean(stepsbyday$stepsbyday, na.rm=TRUE)
median(stepsbyday$stepsbyday, na.rm=TRUE)
```


## What is the average daily activity pattern?
```{r average, echo=TRUE}
library("dplyr")
library("ggplot2")
#calculate average number of steps taken over the 5-minute intervals 
stepsbyday<-activity %>% select (steps, date, interval) %>% group_by(interval) %>% summarize(stepsbyday=mean(steps, na.rm=TRUE))
#graph the plot
ggplot(stepsbyday, aes(x=interval, y=stepsbyday))+ geom_line()+labs(title="avg.steps daily in intervals")
#calculate interval that has max number of steps 
stepsbyday[stepsbyday$stepsbyday==max(stepsbyday$stepsbyday),]
```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
